---
format:
  html:
    include-in-header:
      text: |
        <script src="https://d3js.org/d3.v7.min.js"></script>
page-layout: full
---
::: {.hero}

<div class="hero-image"></div>

<div class="hero-content">

# Who Drives Bycatch?

## Understanding individual variation in manta and devil ray bycatch in the eastern Pacific tuna fishery
Photo credit: TUNACONS

</div>

:::
\



## Catch trends
Manta and Devil rays (mobulids) are captured as bycatch by the EPO tuna purse seine fishery. Catch rates have declined on average over time, possibly due to changes in fishing practices, reporting requirements, and/or declining populations. 
```{r}
#| echo: false

#Reading in data
summary_data <- read.csv("Data/Mob_year_sum.csv")

mobulid_json <- jsonlite::toJSON(
  summary_data,
  dataframe = "rows",
  auto_unbox = TRUE
)
```

<div id="mobulid-scroll-chart"> <svg id="mobulid-svg" viewBox="0 0 800 450"></svg> </div>

```{r}
#| echo: false
#| results: asis
cat(
  '<script>
  const mobulidData = ',
  mobulid_json,
  ';
  </script>'
)
```


<script>

console.log("D3 loaded:", typeof d3);
console.log("Mobulid data:", mobulidData);

const svg = d3.select("#mobulid-svg");

const width = 800;
const height = 450;

const margin = {
  top: 40,
  right: 40,
  bottom: 60,
  left: 70
};

const xScale = d3.scaleLinear()
  .domain([2005, 2025])
  .range([margin.left, width - margin.right]);

const yScale = d3.scaleLinear()
  .domain([0, d3.max(mobulidData, d => d.total_rays)])
  .range([height - margin.bottom, margin.top]);

svg.append("g")
  .attr("transform", `translate(0,${height - margin.bottom})`)
  .call(d3.axisBottom(xScale).tickFormat(d3.format("d")));

svg.append("g")
  .attr("transform", `translate(${margin.left},0)`)
  .call(d3.axisLeft(yScale));

svg.append("text")
  .attr("x", width / 2)
  .attr("y", height - 10)
  .attr("text-anchor", "middle")
  .text("Year");

svg.append("text")
  .attr("transform", "rotate(-90)")
  .attr("x", -(height / 2))
  .attr("y", 20)
  .attr("text-anchor", "middle")
  .text("Total mobulids captured");

const line = d3.line()
  .x(d => xScale(d.Year))
  .y(d => yScale(d.total_rays));

const path = svg.append("path")
  .datum(mobulidData)
  .attr("d", line)
  .attr("fill", "none")
  .attr("stroke", "#2F6F7E")
  .attr("stroke-width", 3);

const length = path.node().getTotalLength();

path
  .attr("stroke-dasharray", length)
  .attr("stroke-dashoffset", length);

function updateChart() {

  const rect = document
    .getElementById("mobulid-svg")
    .getBoundingClientRect();

  const windowHeight = window.innerHeight;

  const start = windowHeight * 0.8;
  const end = windowHeight * 0.2;

  let progress =
    (start - rect.top) / (start - end);

  progress = Math.max(0, Math.min(1, progress));

  path.attr(
    "stroke-dashoffset",
    length * (1 - progress)
  );
}

window.addEventListener("scroll", updateChart);

updateChart();

</script>